# Cleaning the data

Although there is plenty to clean in this data, the most critical pieces currently are ensuring that the
species codes are correct (four characters, all capitalized, no extra spaces, codes are correct for species) and to make sure that any instances where distance and species code data were switched are switched to the correct columns.

## Make a vector of "cleaned" data

First, we need a vector of all of the correct species codes, so we need to know what distinct species were seen, both correct and incorrect:

```{r eval=FALSE}
sobs_obs %>% 
  distinct(species_code)
```

Then, we build the vector of the correct species codes:

```{r eval=FALSE}
corrected_sobs_obs <- data.frame(species_code = c("VESP",
                                   "WEME",
                                   "HOLA",
                                   "ROWR",
                                   "NOBI",
                                   "CORA",
                                   "LASP",
                                   "CONI",
                                   "MODO",
                                   "UNSP",
                                   "CHSP",
                                   "BBMA",
                                   "BUOW",
                                   "PIJA",
                                   "SWHA",
                                   "BTSP",
                                   "BRSP",
                                   "GTTO",
                                   "SATH",
                                   "SPTO",
                                   "COGR",
                                   "SABS",
                                   "EUCD",
                                   "GRAP",
                                   "RTHA",
                                   "LAZB",
                                   "LOSH",
                                   "AMRO",
                                   "UNBI",
                                   "GRFL",
                                   "CHUK",
                                   "HETH",
                                   "PISI",
                                   "NOFL",
                                   "ATFL",
                                   "MOBL",
                                   "RTHU",
                                   "NOHA",
                                   "SACR",
                                   "UNFL",
                                   "WITU",
                                   "AMCR",
                                   "WCSP",
                                   "BHCO",
                                   "UNHA",
                                   "SAPH",
                                   "LBCU",
                                   "SAVS",
                                   "NRWS",
                                   "COPO",
                                   "SCOR",
                                   "HOFI",
                                   "GRSG",
                                   "EUST",
                                   "OCWA",
                                   "HOWR",
                                   "TUVU",
                                   "BRBL",
                                   "PEFA",
                                   "NOMO",
                                   "BLPH",
                                   "INBU",
                                   "AMKE",
                                   "KILL",
                                   "WEKI",
                                   "UNOW",
                                   "YEWA",
                                   "BGGN",
                                   "WAVI",
                                   "BUSH",
                                   "AMGO",
                                   "YBCH",
                                   "BCCH",
                                   "BARS",
                                   "TRES",
                                   "BHGR",
                                   "DOWO",
                                   "BUOR",
                                   "BEWR",
                                   "MOCH",
                                   "GRSP",
                                   "COHA",
                                   "RUGR",
                                   "WTSP",
                                   "RNPH",
                                   "AMPI",
                                   "GRYE",
                                   "CAFI",
                                   "PLVI",
                                   "GWTE",
                                   "GBHE",
                                   "BLJA")
)
```

## Clean the data 

To account for all of the different possibilities that need to be condensed into the correct species codes, we need a long mutate:

```{r eval=FALSE}
cleaned_data <- sobs_obs %>% 
  mutate(clean_species_code = case_when(
    species_code %in% c("Vesp", 
                        "VeSP",
                        "Vesp ",
                        "VESP ",
                        "vesp",
                        "VESp",
                        "VESO",
                        "VESP",
                        "117",
                        "324",
                        "19"
                        ) ~ "VESP",
    species_code %in% c("Weme ",
                        "Weme",
                        "WEmE",
                        "WEME ",
                        "WEMe",
                        "WEMA",
                        "MEWE",
                        "Mela",
                        "MeLa",
                        "MEME",
                        "WEME",
                        "132",
                        "222",
                        "60"
                        ) ~ "WEME",
    species_code %in% c("Hola",
                        "HOLA ",
                        "HOlA",
                        "13 HOLA's",
                        "Hola ",
                        "HOLa",
                        "Horned Lark",
                        "HOL",
                        "HOLA",
                        "107"
                        ) ~ "HOLA",
    species_code %in% c("Rowr ",
                        "Rowr",
                        "ROWR?",
                        "ROWR ",
                        "ROWR"
                        ) ~ "ROWR",
    species_code %in% c("NoBI",
                        "NoBi",
                        "NOBI") ~ "NOBI",
    species_code %in% c("CoRa",
                        "Cora",
                        "CoRa ",
                        "CORA ",
                        "CORA"
                        ) ~ "CORA",
    species_code %in% c("Lasp",
                        "LASP ",
                        "LASL",
                        "LASP"
                        ) ~ "LASP",
    species_code %in% c("CoNI",
                        "CONH",
                        "Night Hawk",
                        "Niha ",
                        "Niha",
                        "NIHA",
                        "CONI",
                        "152"
                        ) ~ "CONI",
    species_code %in% c("MoDo",
                        "Modo",
                        "MODo",
                        "MODo ",
                        "MODO ",
                        "MOSO",
                        "MODO"
                        ) ~ "MODO",
    species_code %in% c("Unsp ",
                        "Unsp",
                        "UNSP ",
                        "UNSP"
                        ) ~ "UNSP",
    species_code %in% c("Chsp ",
                        "Chsp",
                        "CHSP"
                        ) ~ "CHSP",
    species_code %in% c("Bbma ",
                        "Bbma",
                        "BLMA",
                        "BBMP",
                        "BTMJ",
                        "BBMA"
                        ) ~ "BBMA",
    species_code %in% c("BAOW",
                        "BUOW"
                        ) ~ "BUOW",
    species_code %in% c("Pija",
                        "PIJA"
                        ) ~ "PIJA",
    species_code %in% c("Swha ",
                        "SWHA ",
                        "SWHA"
                        ) ~ "SWHA",
    species_code %in% c("BLSP",
                        "BTSP"
                        ) ~ "BTSP",
    species_code %in% c("Brsp ",
                        "Brsp",
                        "BRSP ",
                        "Brs",
                        "Brap",
                        "BRAP",
                        " BRSP"
                        ) ~ "BRSP",
    species_code %in% c("Gtto ",
                        "Gtto",
                        "gtto",
                        "GTTO ",
                        "GRTO",
                        " GTTO",
                        "20"
                        ) ~ "GTTO",
    species_code %in% c("SATH ",
                        "Sath",
                        "SATH",
                        "151"
                        ) ~ "SATH",
    species_code %in% c("Spto ",
                        "Spto",
                        "Spotted towhee",
                        "SPTO"
                        ) ~ "SPTO",
    species_code %in% c("Sabs ",
                        "Sabs",
                        "SABS ",
                        "Sags",
                        "SASP",
                        "SASP ",
                        "SBSP",
                        "SABS"
                        ) ~ "SABS",
    species_code %in% c("GRPA",
                        "GRAP"
                        ) ~ "GRAP",
    species_code %in% c("Rtha ",
                        "Rtha",
                        "RTHA"
                        ) ~ "RTHA",
    species_code %in% c("Lazb ",
                        "LABU",
                        "Labu ",
                        "Lazuli bunting",
                        "LAZB"
                        ) ~ "LAZB",
    species_code %in% c("LHSH",
                        "LOSH"
                        ) ~ "LOSH",
    species_code %in% c("UN",
                        "UNKNOWN",
                        "UNBI"
                        ) ~ "UNBI",
    species_code == " WCSP" ~ "WCSP",
    species_code %in% c("AMRA",
                        "AMRO"
                        ) ~ "AMRO",
    species_code %in% c("Gray FlyCatcher",
                        "Gray FlyCatcher ",
                        "GRFL"
                        ) ~ "GRFL",
    species_code %in% c("chukar",
                        "chukar ",
                        "chukar?",
                        "Chukar",
                        "Chuk",
                        "Chuk ",
                        "Chukar?",
                        "CHUCK"
                        ) ~ "CHUK",
    species_code %in% c("Nofl ",
                        "Nofl",
                        "Northern flicker"
                        ) ~ "NOFL",
    species_code == "ATFC" ~ "ATFL",
    species_code == "Northern harrier" ~ "NOHA",
    species_code %in% c("Sacr ",
                        "Sacr",
                        "SAND"
                        ) ~ "SACR",
    species_code %in% c("Flycatcher",
                        "Flycatcher "
                        ) ~ "UNFL",
    species_code == "Turkey" ~ "WITU",
    species_code %in% c("Bhco ",
                        "BRCO",
                        "BHCB"
                        ) ~ "BHCO",
    species_code %in% c("Lbcu ",
                        "Lbcu",
                        "LOCU"
                        ) ~ "LBCU",
    species_code %in% c("SAGR",
                        "Greater sage grouse "
                        ) ~ "GRSG",
    species_code == "Howr " ~ "HOWR",
    species_code %in% c("Tuvu ",
                        "Tuvu"
                        ) ~ "TUVU",
    species_code %in% c("Brbl ",
                        "Brbl",
                        "BRBB",
                        "10 BRBB's"
                        ) ~ "BRBL",
    species_code == "Spotted falcon " ~ "PEFA",
    species_code %in% c("Amke ",
                        "Amke",
                        "MAKE"
                        ) ~ "AMKE",
    species_code == "KIDE" ~ "KILL",
    species_code == "Unknown owl" ~ "UNOW",
    species_code == "BLGN" ~ "BGGN",
    species_code == "bushtit" ~ "BUSH",
    species_code == "BASW" ~ "BARS",
    species_code == "TRSW" ~ "TRES",
    species_code == "GHSP" ~ "GRSP",
    species_code == "Coha" ~ "COHA",
    TRUE ~ as.character(species_code)
  ))
```

With this new column, we now have almost all of the correct species information. The only thing left is to correctly place the instances were distance and species code data got mixed up:

```{r eval=FALSE}
cleaned_data[251, 3] = "WEME"
cleaned_data[251, 4] = 132
cleaned_data[292, 3] = "VESP"
cleaned_data[302, 3] = "WEME"
cleaned_data[302, 4] = 222
cleaned_data[2002, 3] = "VESP"
cleaned_data[2002, 4] = 53
cleaned_data[2197, 3] = "VESP"
cleaned_data[2917, 4] = 39
cleaned_data[2197, 4] = 324
cleaned_data[2426, 3] = "VESP"
cleaned_data[2426, 4] = 19
cleaned_data[3116, 3] = "GTTO"
cleaned_data[3116, 4] = 20
cleaned_data[4386, 3] = "SPTO"
cleaned_data[4386, 4] = 53
cleaned_data[11100, 3] = "HOLA"
cleaned_data[11100, 4] = 107
cleaned_data[11104, 3] = "WEME"
cleaned_data[11104, 4] = 60
cleaned_data[11207, 3] = "CONI"
cleaned_data[11207, 4] = 152
cleaned_data[11584, 3] = "SATH"
cleaned_data[11584, 4] = 151
```


Now the species code data is completely cleaned, and the distance and species code data is located in the correct columns throughout the data frame.